### 2014-12-19 Studies/2013_PG_County/Report/PG_Co_Fire.tex
PG Co: add figures for flow path and update report

\begin{table}[!ht]
\caption[Fire incident and simulation event timeline]
{Fire incident and simulation event timeline}
\begin{tabular}{ccl}
\toprule
Incident Time  &  Simulation  &  Fire Behavior / Fireground Operation                                \\
               &  Time        &                                                                      \\
{(hh:mm:ss)}   &  {(s)}       &                                                                      \\
\midrule
21:08:26       &              &  First 911 call is made.                                             \\
               &              &                                                                      \\
21:11:03       &              &  Dispatch for a``box'' alarm assignment- four engine companies       \\
               &              &  two truck companies, one rescue squad, and a commanding officer.  \\
               &              &                                                                      \\
21:21:20       &  0           &  FDS simulation begins.                                              \\
               &              &                                                                      \\
21:12:55       &              &  FDFirst unit (Engine 807B) arrived on-scene.                        \\
               &              &                                                                      \\
21:13:44       &              &  Volunteer Chief 809A establishes incident command (IC).             \\
               &              &                                                                      \\
21:13:56       &              &  Truck 809 arrived on-scene.                                         \\
               &              &                                                                      \\
21:14:00       &  100         &  Front door opened.                                                  \\
               &              &                                                                      \\
21:14:42       &              &  Truck 809 enters structure.                                         \\
               &              &                                                                      \\
21:15:47       &  207         &  Front door closed over.                                             \\
               &              &  Basement door opened and water applied to fire.                     \\
               &              &                                                                      \\
21:15:51       &  211         &  Lower pane of front egress window broke.                            \\
               &              &                                                                      \\
21:16:01       &  221         &  Top pane of front egress window broke.                              \\
               &              &                                                                      \\
21:16:16       &              &  Truck 809 officer bailout through window.                           \\
               &              &                                                                      \\
21:16:47       &  267         &  Front door opened.                                                  \\
               &              &                                                                      \\
21:17:15       &              &  Truck 809 forcible entry firefighter rescued.                       \\
               &              &                                                                      \\
\bottomrule
\end{tabular}
\footnotesize
\\ For specific information regarding ventilation areas used in the simulation, refer to Table~\ref{tab:vents}.
\normalsize
\label{tab:combined_timeline}
\end{table}

@@

Analysis of the temperatures from the simulation focuses on the high-hazard area of the structure: the first floor area between the front door and the top of the interior stairwell. The temperature contours shown in Fig.~\ref{fig:temperature_slice_99s} indicate the gas temperatures within the structure 1~s before the front door was opened 0.91~m (3~ft) above the floor. The overlayed arrows represrent the gas flow: the length of the lines represents speeds and the direction of the arrow represents the flow direction. Prior to the door opening, the open living area and kitchen on the first floor were approximately 60~$^{\circ}$C (140~$^{\circ}$F) while the peak temperatures were concentrated near the interior stairwell, 105~$^{\circ}$C (220~$^{\circ}$F). Note that there is very minimal flow throughout the first floor. Figures~\ref{fig:temperature_slice_101s} and \ref{fig:temperature_slice_105s} show gas temperature contours at 0.91~m (3~ft) above the floor, 1~s and 5~s after the front door was opened.

@@

After the front door is opened, hot fire gases from the basement begin to flow through the interior stairwall into the first floor. Figure~\ref{fig:temperature_slice_101s}, 1~s after the front door was opened, shows that the temperature contours at the top of the stairwell have increased to approximately 200~$^{\circ}$C (400~$^{\circ}$F) with a hot spot that approaches 425~$^{\circ}$C (800~$^{\circ}$F). Figure~\ref{fig:temperature_slice_105s}, 5~s after the front door was opened, shows more significant increases in temperature. The area of at the top of the stairs is most between 375~$^{\circ}$C (700~$^{\circ}$F) and 425~$^{\circ}$C (800~$^{\circ}$F). The gas temperatures 0.91~m (3~ft) above the floor at the front door have also increased to approximately 230~$^{\circ}$C (450~$^{\circ}$F) and show increased flow velocities.

@@

\begin{figure}[!ht]
\includegraphics[trim = 3in 3in 3in 2in, clip=true, width=.65\textwidth]{../Figures/flow_path_99s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Flow path vectors at front door and top of stairwell 1~s before front door is opened]
{Flow path vectors at front door and top of stairwell 1~s before front door is opened.}
\label{fig:flow_path_99s}
\end{figure}

>>>>

\begin{table}[!ht]
\caption[Fire incident and simulation event timeline]
{Fire incident and simulation event timeline}
\begin{tabular}{ccl}
\toprule
Incident Time  &  Simulation  &  Fire Behavior / Fireground Operation                                \\
               &  Time        &                                                                      \\
{(hh:mm:ss)}   &  {(s)}       &                                                                      \\
\midrule
21:08:26       &              &  First 911 call is made.                                             \\
               &              &                                                                      \\
21:11:03       &              &  Dispatch for a``box'' alarm assignment- four engine companies       \\
               &              &  two truck companies, one rescue squad, and a commanding officer.    \\
               &              &                                                                      \\
21:21:20       &  0           &  FDS simulation begins.                                              \\
               &              &                                                                      \\
21:12:55       &              &  First unit (Engine 807B) arrived on-scene.                          \\
               &              &                                                                      \\
21:13:44       &              &  Volunteer Chief 809A establishes incident command (IC).             \\
               &              &                                                                      \\
21:13:56       &              &  Truck 809 arrived on-scene.                                         \\
               &              &                                                                      \\
21:14:00       &  100         &  Front door opened.                                                  \\
               &              &                                                                      \\
21:14:42       &              &  Truck 809 enters structure.                                         \\
               &              &                                                                      \\
21:15:47       &  207         &  Front door closed over.                                             \\
               &              &  Basement door opened and water applied to fire.                     \\
               &              &                                                                      \\
21:15:51       &  211         &  Lower pane of front egress window broke.                            \\
               &              &                                                                      \\
21:16:01       &  221         &  Top pane of front egress window broke.                              \\
               &              &                                                                      \\
21:16:16       &              &  Truck 809 officer bailout through window.                           \\
               &              &                                                                      \\
21:16:47       &  267         &  Front door opened.                                                  \\
               &              &                                                                      \\
21:17:15       &              &  Truck 809 forcible entry firefighter rescued.                       \\
               &              &                                                                      \\
\bottomrule
\end{tabular}
\footnotesize
\\ For specific information regarding ventilation areas used in the simulation, refer to Table~\ref{tab:vents}.
\normalsize
\label{tab:combined_timeline}
\end{table}

@@

Analysis of the temperatures from the simulation focuses on the high-hazard area of the structure: the first floor area between the front door and the top of the interior stairwell. The temperature contours shown in Fig.~\ref{fig:temperature_slice_99s} indicate the gas temperatures within the structure 1~s before the front door was opened 0.91~m (3~ft) above the floor. The overlaid arrows represent the gas flow: the length of the lines represents speeds and the direction of the arrow represents the flow direction. Prior to the door opening, the open living area and kitchen on the first floor were approximately 60~$^{\circ}$C (140~$^{\circ}$F) while the peak temperatures were concentrated near the interior stairwell, 105~$^{\circ}$C (220~$^{\circ}$F). Note that there is very minimal flow throughout the first floor. Figures~\ref{fig:temperature_slice_101s} and \ref{fig:temperature_slice_105s} show gas temperature contours at 0.91~m (3~ft) above the floor, 1~s and 5~s after the front door was opened.

@@

After the front door is opened, hot fire gases from the basement begin to flow through the interior stairwell into the first floor. Figure~\ref{fig:temperature_slice_101s}, 1~s after the front door was opened, shows that the temperature contours at the top of the stairwell have increased to approximately 200~$^{\circ}$C (392~$^{\circ}$F) with a hot spot that approaches 425~$^{\circ}$C (800~$^{\circ}$F). Figure~\ref{fig:temperature_slice_105s}, 5~s after the front door was opened, shows more significant increases in temperature. The area of at the top of the stairs is most between 375~$^{\circ}$C (700~$^{\circ}$F) and 425~$^{\circ}$C (800~$^{\circ}$F). The gas temperatures 0.91~m (3~ft) above the floor at the front door have also increased to approximately 230~$^{\circ}$C (450~$^{\circ}$F) and show increased flow velocities.

@@

The report from Prince George's County Fire/EMS~\cite{PGCounty2013} indicated that two firefighters from Truck 809 were injured after making entry to the first floor and had to either bail out through a window or be rescued through the front door. From Section~\ref{temp}, the conditions on the first floor were initially tenable. After the front door was opened, however, the simulation results indicate that a high hazard area existed in the first floor that exceeded the conditions of a Class~III exposure (temperatures greater than 260~$^{\circ}$C or 500~$^{\circ}$F)~\cite{Donnelly2006}.

The simulation results indicate that a flow path was established between the open vents (windows and door) on the rear side of the basement and front side of the structure (through the front door of the first floor) after the front door was opened. Strong winds (8.9~m/s (20~mph)) pushed air into the rear side of the basement through window failures and combined with buoyant combustion gases to pressurize the structure. Opening the front door provided a pressure relief; this resulted in a rapid change in the conditions within the structure and the establishment of a flow path. After the front door was opened, the combustion gases at elevated temperature and pressure in the basement flowed towards lower pressure regions outside of the structure on the leeward side (front of structure) via the interior stairwell. The two injured firefighters were located in the flow path on first floor between the top of the interior stairwell and the front door.

To examine the impact of opening the front door on the establishment of a flow path, flow conditions are examined along the centerline of the interior stairwell doorway and at the front door. Figure~\ref{fig:flow_path_99s} shows the flow vectors at the first floor stairwell doorway 1 second before the front door was opened. Note that length and direction of the arrow indicates the magnitude and direction of the flow respectively. The color of the arrow corresponds to the values from the included temperature scales. Therefore, a short blue arrow represents slow moving, cool gas while a long red arrow represents fast moving, high temperature gas. Prior to the door opening, there is minimal flow through the first floor stairwell door and the temperature peak is approximately 125~$^{\circ}$C (257~~$^{\circ}$F) as shown in Figure~\ref{fig:flow_path_99s}. One second after the door opened, Figure~\ref{fig:flow_path_101s}, there is noticeable flow through both the stairwell doorway and through the front door. Simulation temperatures at the stairwell increased to approximately 200~$^{\circ}$C (392~~$^{\circ}$F) and temperatures through the doorway are approximately 100~$^{\circ}$C (212~~$^{\circ}$F). These values are also shown in planar contours 1~m above the floor in Figs.~\ref{fig:velocity_slices} and \ref{fig:temperature_slice_101s} for velocity and temperature, respectively.

\begin{figure}[!ht]
\includegraphics[trim = 2.5in 0in 4in 0in, clip=true, width=.65\textwidth]{../Figures/flow_vector_99s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Flow path vectors at front door and top of stairwell 1~s before front door was opened]
{Flow path vectors at front door and top of stairwell 1~s before front door was opened.}
\label{fig:flow_path_99s}
\end{figure}

\begin{figure}[!ht]
\includegraphics[trim = 2.5in 0in 4in 0in, clip=true, width=.65\textwidth]{../Figures/flow_vector_101s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Flow path vectors at front door and top of stairwell 1~s after front door was opened]
{Flow path vectors at front door and top of stairwell 1~s after front door was opened.}
\label{fig:flow_path_101s} 
\end{figure}

At the estimated time of firefighter entry to the first floor (142~s into the simulation; 21:14:42 according to Table~\ref{tab:combined_timeline}), the simulated flow magnitudes through the first floor (stairwell and front doors) remained similar to the conditions just after opening the front door. The modeln temperatures increased through the stairwell doorway to between 400~$^{\circ}$C and 425~$^{\circ}$C (750-800~$^{\circ}$F) and through the front door to approximately 375~$^{\circ}$C (700~$^{\circ}$F). From the approximate timeline (Table~\ref{tab:combined_timeline}), the front door was forced closed due to airflow pushing against the front door 207~s into the simulation. Figures~\ref{fig:flow_path_206s} and \ref{fig:flow_path_207s} show the conditions 1~s prior to and after the front door closed. The figures show that while the closed door stopped significant flow through the stairwell door and onto the first floor, temperature hazard for the firefighters (temperatures at or above 400~$^{\circ}$C or 750~$^{\circ}$F) remained.

\begin{figure}[!ht]
\includegraphics[trim = 2.5in 0in 4in 0in, clip=true, width=.65\textwidth]{../Figures/flow_vector_142s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Flow path vectors at front door and top of stairwell at estimated time of first floor entry]
{Flow path vectors at front door and top of stairwell 42~s after front door was opened, the estimated time of first floor entry.}
\label{fig:flow_path_142s}
\end{figure}

\begin{figure}[!ht]
\includegraphics[trim = 2.5in 0in 4in 0in, clip=true, width=.65\textwidth]{../Figures/flow_vector_206s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Flow path vectors at front door and top of stairwell 1~s before front door was closed]
{Flow path vectors at front door and top of stairwell 1~s before front door was closed.}
\label{fig:flow_path_206s}
\end{figure}

\begin{figure}[!ht]
\includegraphics[trim = 2.5in 0in 4in 0in, clip=true, width=.65\textwidth]{../Figures/flow_vector_207s}
\input{../Scripts/colorbar_metric_temp}
\input{../Scripts/colorbar_std_temp}
\caption[Flow path vectors at front door and top of stairwell 1~s after front door was closed]
{Flow path vectors at front door and top of stairwell 1~s after front door was closed.}
\label{fig:flow_path_207s}
\end{figure}
